---
title: "Updating Data to be used in SRA"
author: "Julia Levine & Natalie Bryce"
date: "`r format(Sys.time(), '%d %B, %Y')`"
output: 
  pdf_document:
    toc: true
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, warning = F, message = F)
library(data.table)
library(knitr)
library(zoo)
library(ggplot2)
library(dplyr)
#rm(list = ls())
# the year of the data we are adding in
update_years <- c(2022)
first.pass <- F
setwd("/Users/nmbryce/Documents/R/SRA/2023-Statistical-Risk-Assessment-github/2.Make-new-base-adata")
base_data = "output/prepared2022predictors-no2022-2023-10-26-2.csv"
```




\newpage


# Load in old data

The following code reads in the old data. It then takes the most recent data for each country and replicates it for the years that we are updating -- this let's us easily keep the time-invariant values, like region. I then set aside the old data, and update the data that will be appended.

```{r}
# read in data from the previous years , called dat 

  dat <- fread(base_data)

# carry forward all data from the most recent year

  carry_from <- max(dat$year)
  
  # this only applies if updating multiple years simultaneously
  # it carries forward the most recent data for each of the years to be updated
  dat_new <- rbindlist(rep(list(dat[year == carry_from]), times = length(update_years)))
  
  # relabel the years
  
  dat_new$year <- rep(update_years, each = nrow(dat_new))

# rename old data and leave it alone until you've updated all 
# variables except for the lag/leads
  
  dat_old <- dat
  rm(dat)
  

```



# Update 

## Define functions

### carry_forward

The following is a function to carry forward recent values of a variable for countries where it's NA

```{r}
  carry_forward <- function(variable, 
                            carry_to = update_years, 
                            data = dat, 
                            carry_from = (update_years - 1)){
    na_var <- data[year == carry_to & is.na(get(variable))]$ccode
    
    cat(paste0("Missing values for ", length(na_var), " countries."))

    carry_var <- subset(data[year == carry_from & ccode %in% na_var], 
                        select = c("ccode", variable))
    cat(paste0("\nFilling in values for ", 
               nrow(carry_var[!is.na(get(variable))]), " countries."))

    for(i in carry_var[!is.na(get(variable))]$ccode){
      data[year == carry_to & ccode == i, 
          (variable) := carry_var[ccode == i, get(variable)]]
      data[year == carry_to & ccode == i, (paste(variable, "carry_from", sep = "_")) := carry_from]
    }
  }
```

### merge_dat

The following is a function that removes the variables that were carried forward from the base data, and then merges in the updated variables from the appropriate year.

```{r}
merge_dat <- function(new_data, base_data = dat_new, 
                      merge_new_by = c("ccode", "year"), 
                      merge_base_by = c("ccode", "year")){
  new_data_vars <- colnames(new_data)[!colnames(new_data) %in% merge_new_by]
  dat1 <- subset(base_data, 
                 select = colnames(base_data)[!(colnames(base_data) %in% 
                                                  new_data_vars)])
  dat1 <- merge(dat1, new_data, 
                by.x = merge_base_by, 
                by.y = merge_new_by, all.x = TRUE)
  dat1
}
```



## Country age

This chunk updates country age by adding the difference between the updated years and the year the data was carried from to the country age variable. 

```{r}
# update country age

  diff <- update_years - carry_from
  
  for(i in 1:length(update_years)){
    dat_new$countryage[dat_new$year == update_years[i]] <- 
    dat_new$countryage[dat_new$year == update_years[i]] + diff[i]
  }

# update logged country age variable

  dat_new$countryage.ln <- log(dat_new$countryage + 1)
```



## VDEM

The following code reads in the most recent V-DEM data, selecting the relevant variables and renaming them. 

```{r, eval = FALSE}


  vdemfull <- fread("input/V-Dem-CY-Full+Others-v13.csv")
  
# subset to just use these variables 
#note v2mecenefm_ord and e_peinfmor was added for the 2023 update


  vdem_vars <- c("v2cldmovem_ord", 
                 "v2cldmovew_ord", 
                 "v2clkill_ord",
                 "v2clsocgrp_ord", 
                 "v2clrgunev_ord", 
                 "v2csreprss_ord", 
                 "v2pepwrsoc_ord", 
                 "v2elrstrct",
                 "v2psparban_ord", 
                 "v2psoppaut_ord", 
                 "v2jureform_ord",
                 "v2clrelig_ord", 
                 "v2xcl_disc", 
                 "v2pepwrses_ord", 
                 "v2pepwrses",
                 "e_migdpgro", 
                 "e_mipopula", 
                 "e_cow_exports", 
                 "e_cow_imports", 
                 "e_migdppc", 
                 "v2csgender_ord",
                 "v2pepwrgen_ord", 
                 "v2mecenefm_ord",
                 "e_peinfmor",
                 "e_pop",
                 "e_gdppc")

  
  keep <- c("COWcode", "country_name", "year", vdem_vars)

# function that renames and formats variables, 
# takes dataset and vector of variables to keep as inputs
  
  source("helper scripts/format_vdem.R")
  
# function that maps COW codes onto PITF country codes, 
# contained in format_vdem 
  
  source("helper scripts/cowtopitf2018.R")
  

# format and save vdem

  vdem <- format_vdem(dat = vdemfull, keep = keep)

save(vdem, file = "input/vdem.Rdata")
```

```{r}
# read in formatted VDEM variables for the update years 
  
  load("input/vdem.Rdata")

# limit to the update years
  
  vdem <- vdem[year %in% update_years]

# count missing values
# check which variables are missing values
  
  missing <- apply(vdem[COWcode %in% dat_new$ccode], 2, 
                   function(x) sum(is.na(x)))
  kable(missing, col.names = "NA count")

# check which variables are missing for all countries, indicating that they
# have not been updated yet, and are potentially discontinued
  
  all.missing <- names(missing)[missing == nrow(vdem)]
  if(length(all.missing) > 0){
      cat(paste0("The following variables have not been updated as of ", 
            Sys.Date(),":\n", paste(all.missing, collapse = "\n")))
  }

  
# this ensures that we don't replace old country names
# VDEM sometimes changes them
  
  vdem$country_name <- NULL
  
  #rename Vdem's COW code column
  
  colnames(vdem)[colnames(vdem) == 'COWcode'] <- 'ccode'

# update V-DEM data
  
  dat_new <- merge_dat(new_data = vdem)
  

```


## WDI

This chunk pulls select variables from the WDI API.


```{r, eval = FALSE}
# In June 2020 WDI decommissioned V1 of the API
# so you may need to re-install the WDI package.  
# install.packages("devtools")
  # devtools::install_github("vincentarelbundock/WDI")
  library(WDI)

# list indicators we want to pull
  
  wdilist <- c("NE.TRD.GNFS.ZS",     # Trade (% of GDP)
               "NY.GDP.PCAP.KD",     # GDP per capita (constant 2000 US$)
               "NY.GDP.MKTP.KD.ZG",  # GDP growth (annual %)
               "SP.POP.TOTL",        # Population, total
               "SP.DYN.IMRT.IN"      # Infant mortality rate
               
      )

  # confirm that all of the old indicator names are in use 
  
    (check.names <- rbindlist(lapply(wdilist, function(x) {
      # search the indicator list for each of the variables in wdilist
       res <- WDIsearch(x, field = "indicator") 
       # take the first match (drops those with extra letters on the end)
       if(length(res) > 2){
         res <- res[1, ]
       } 
       data.table("indicator" = res[1], "name" = res[2])})))

# Extract latest version of desired variables from WDI 
# also pull 5 years from before so that we can carry forward
# more recent values for variables like infant mortality that are slow to update
  
  wdi <- WDI(country="all", indicator=wdilist, extra=FALSE, 
             start=(min(update_years) - 5))

#add COW codes for merging, remove country column to avoid merging that in later 
  library(countrycode)
  
  wdi <- wdi %>% 
  mutate(ccode = countrycode(iso3c, origin = "iso3c", destination = "cown"))
  
  wdi <- wdi %>% 
  mutate(ccode = ifelse(country == "Kosovo",347,ccode),
         ccode = ifelse(country == "Serbia",345,ccode))


#See which cowcodes don't match - we see that these are all regions or small states that we don't include in our dataset

wdi %>% 
  anti_join(dat_new, by = c("ccode")) %>% 
  select(year, country)

#checking for NAs, the only issues are Kosovo - I change it to 347 - and Serbia - changed to 345

wdi %>% 
  filter(is.na(ccode))

#remove country name
wdi <- wdi %>% 
  select(-country)


# Reorder for easier review
  
  wdi <- wdi[order(wdi$ccode, wdi$year),]

# Rename variables-- add a "new" to indicate these are newly brought in from wdi
# to avoid reusing names already in the old EWP data
  
  setDT(wdi)
  wdi_cols <- c("wdi.trade.new",
             "wdi.gdppc.new",
             "wdi.gdppcgrow.new",
             "wdi.popsize.new", 
             "wdi.imrate.new")
  setnames(wdi, 
           c("NE.TRD.GNFS.ZS",    
             "NY.GDP.PCAP.KD",    
             "NY.GDP.MKTP.KD.ZG", 
             "SP.POP.TOTL",       
             "SP.DYN.IMRT.IN"), 
           wdi_cols)


# save to the input folder
  if(first.pass == T){dir.create("input/temp/wdi")}
  fwrite(wdi, paste0("input/temp/wdi/wdipulled-", Sys.Date(), ".csv"))
```





This chunk reads in the latest pull from the WDI API, counts the missing values, and merges it into the main dataset.

```{r}
  
 wdi <- fread("input/temp/wdi/wdipulled-2023-10-26.csv")

#create "new.2" versions so that they merge correctly with those vars as well
wdi <- wdi %>% 
  mutate(wdi.trade.new.2 = wdi.trade.new,
         wdi.gdppc.new.2 = wdi.gdppc.new,
         wdi.gdppcgrow.new.2 = wdi.gdppcgrow.new,
         wdi.imrate.new.2 = wdi.imrate.new)

new.wdi <- wdi[year %in% update_years]

  
# check which variables are missing values
  
  wdi_cols <- c("wdi.trade.new.2",
             "wdi.gdppc.new.2",
             "wdi.gdppcgrow.new.2",
             "wdi.popsize.new", 
             "wdi.imrate.new.2")
  missing = melt(new.wdi[, lapply(.SD, function(x) sum(is.na(x))), .SDcols = wdi_cols])

  kable(missing, col.names = c("Variable", "NA count"))

# check which variables are missing for all countries, indicating that they
# have not been updated yet, but potentially will be in the future
  
  all.missing <- missing[value == length(unique(wdi$sftgcode))]$variable
  cat(paste("\n", all.missing, "not updated as of", Sys.Date()))

  #Hold on carrying forward the last non-missing value of infant mortality
  # Carry forward last non-missing values of tradeshare and gdp as they were mostly updated in 2018 and the earliest values from CIA factbook are 2017

  carry_forward(variable = "wdi.trade.new", carry_from = update_years - 1, data = wdi)
  carry_forward(variable = "wdi.trade.new", carry_from = update_years - 2, data = wdi)
  carry_forward(variable = "wdi.trade.new", carry_from = update_years - 3, data = wdi)
  
  carry_forward(variable = "wdi.gdppc.new", carry_from = update_years - 1, data = wdi)
  carry_forward(variable = "wdi.gdppc.new", carry_from = update_years - 2, data = wdi)
  carry_forward(variable = "wdi.gdppc.new", carry_from = update_years - 3, data = wdi)

  
  missing = melt(new.wdi[, lapply(.SD, function(x) sum(is.na(x))), .SDcols = wdi_cols])
  kable(missing, col.names = c("Variable", "NA count"))
  
# Merge it in to main data
  
  
  dat_new <- merge_dat(new_data = wdi[year %in% update_years])
  dat_new[, imr.sqrt.new := sqrt(wdi.imrate.new)]
  dat_new[, imr.sqrt.new.2 := sqrt(wdi.imrate.new.2)]
  dat_new[, wdi.trade.ln.new := log(wdi.trade.new)]
  dat_new[, wdi.trade.ln.new.2 := log(wdi.trade.new.2)]
  dat_new[,wdi.popsize.ln := log(wdi.popsize.new)]

  
```



## UCDP battle-related deaths 


The following code reads in the [UCDP data](https://ucdp.uu.se/downloads/) and changes country names to match. 
Below, I sum the number of battledeaths for each country in the update years, and merge this onto dat_new. 


```{r, eval = F}
# read in UCDP data
  
  ucdp <- fread("input/BattleDeaths_v23_1.csv")

# making sure we have the same variables as in 18.1 version. 
# colnames for 18.1 are uppercase, use tolower when checking

  colnames(ucdp) <- gsub("_", "", colnames(ucdp))

# limit to the update years and conflict type >= 3
  
  ucdp <- ucdp[year %in% update_years & typeofconflict >= 3]

# change country names to match
  
  diff_loc <- setdiff(unique(ucdp$locationinc), unique(dat_new$country))
  
  # drops the locations that involve multiple countries
  
    diff_loc <- diff_loc[grepl(", ", diff_loc) == F]
  
  # change names
    
    ucdp$locationinc[ucdp$locationinc=="Burkina Faso"] = "Burkina Faso (Upper Volta)"
    ucdp$locationinc[ucdp$locationinc=="Yemen (North Yemen)"] = "Yemen (unified)"
    ucdp$locationinc[ucdp$locationinc=="DR Congo (Zaire)"] = "Congo, Democratic Republic of (Zaire)"
    ucdp$locationinc[ucdp$locationinc=="Turkey"] = "Turkey (Ottoman Empire)"


    
    if(length(setdiff(unique(ucdp$locationinc), 
                      unique(dat_new$country)))>0){
      stop("Different country names")}
    
    
# group by year/country and sum battledeaths
    
  bd <- ucdp[, .("battledeaths" = sum(bdbest)), by = c("locationinc", "year")]
  
# save battledeath data
  
  fwrite(bd, "input/temp/battledeaths.csv")
```



I then read in the battledeaths for each country and merge it into the new data, replacing missing values with zeroes and updating the logged version of this variable.

```{r}
# read in battledeaths data

  bd <- fread("input/temp/battledeaths.csv")

# merge battledeaths into main data
    
  dat_new <- merge_dat(new_dat = bd,
                       merge_new_by = c("locationinc", "year"), 
                       merge_base_by = c("country", "year"))

    
  
# fill in zeroes and update logged variable
  
  setDT(dat_new)
  
  dat_new[, battledeaths := ifelse(is.na(battledeaths), 0, battledeaths)]
  dat_new[, battledeaths.ln := log(battledeaths + 1)]
  
```


## Coup Attempts

This code reads in the most updated version of the Powell and Thyne data which is posted on their website. It selects coups in the years to be updated


```{r, eval = FALSE}

# pull coup data from powell and thyne website

  coup_dat <- as.data.table(read.delim("http://www.uky.edu/~clthyn2/coup_data/powell_thyne_coups_final.txt"))

# keep coups in the update years
  
  new_coup <- coup_dat[year %in% update_years]

# where coup == 2 it was a successful coup
# where coup == 1 it was a failed coup
  
  new_coup[, cou.s.d := ifelse(coup == 2, 1, 0)]
  new_coup[, cou.f.d := ifelse(coup == 1, 1, 0)]

# save coup data
  
  fwrite(new_coup, paste0("input/temp/", "new-coups-pulled-", Sys.Date(), ".csv"))
```



This updates the coup variables in dat_new.

```{r}
# read in the coup data

 # temp.files <- list.files("input/temp")
 # coup.files <- grep("coups", temp.files, value = T)
 # latest.pull <- which.max(lapply(coup.files, 
     #                             function(x) {
      #                              as.Date(gsub("coups-pulled-|.csv", "", x))}))
 # new_coup <- fread(paste0("input/temp/", coup.files[latest.pull]))
  
  new_coup <- fread("input/temp/new-coups-pulled-2023-10-07.csv")

# select variables and print
  
  new_coup <- new_coup[, .(cou.s.d = as.integer(sum(cou.s.d)>0), 
               cou.f.d = as.integer(sum(cou.f.d)>0)), by = .(country, year)]
  
  kable(new_coup)
  
######################################################################
# CHECK NAMES BEFORE MERGE
######################################################################
  
  check_names = rbindlist(lapply(seq_len(nrow(new_coup)), function(i){
    dat_new[grepl(new_coup[i]$country, country),
            .(coup_country = new_coup[i]$country, 
              countries_in_sra = country)]
  }))
  check_names[, exact_match := coup_country == countries_in_sra]
  check_names[, has_exact_match := sum(exact_match)>0, by = .(coup_country)]
  check_names = check_names[!(has_exact_match == T & exact_match == F)]
  check_names[has_exact_match == F]
  new_coup[country == "Burkina Faso", country := "Burkina Faso (Upper Volta)"]
  

######################################################################  
  
# merge coup data into dat_new
  
  dat_new <- merge_dat(new_data = new_coup, 
                       merge_new_by = c("country", "year"), 
                       merge_base_by = c("country", "year"))

# fill in missing values with zeroes

  coup_cols <- c("cou.s.d", "cou.f.d")
  dat_new[, (coup_cols) := lapply(.SD, function(x) 
    ifelse(is.na(x), 0, x)), .SD = coup_cols]

# update cou.any variable (1 if there was either a failed or successful coup)
  
  dat_new[, cou.any := ifelse(cou.s.d>0 | cou.f.d>0, 1, 0)]
  
```

#merge in new mass killing data
In a typical year, this would consist of manually making updates to the relevant MK vars based on changes in the prior year. This year, we recoded the DV so below I merge in 2022 values for the relevant MK variables.

```{r}
base_mk_data = "input/basedata2023_13sept2023.csv"

  mk_dat_2022 <- fread(base_mk_data)
  
  #limit to 2022 and relevant columns
  
  mk_dat_2022 <- mk_dat_2022 %>% 
    filter(year == 2022) %>% 
    select(-country,-V1,)
  
  dat_new <- dat_new %>% 
    left_join(mk_dat_2022, by = c("ccode","year")) %>% 
    mutate(datastartyear = ifelse(year == 2022, datastartyear.y,datastartyear.x),
           dataendyear = ifelse(year == 2022, dataendyear.y,dataendyear.x),
           mk_onset_count = ifelse(year == 2022, mk_onset_count.y, mk_onset_count.x),
           mk_offset_count = ifelse(year == 2022, mk_offset_count.y,mk_offset_count.x),
           mk_ongoing_count = ifelse(year == 2022, mk_ongoing_count.y,mk_ongoing_count.x),
           mk_location_ever = ifelse(year == 2022, mk_location_ever.y,mk_location_ever.x),
           mk_responsible_ever = ifelse(year == 2022, mk_responsible_ever.y, mk_responsible_ever.x),
           mk_ongoing_widescope = ifelse(year == 2022, mk_ongoing_widescope.y, mk_ongoing_widescope.x),
           mk_ongoing_code = ifelse(year == 2022, mk_ongoing_code.y, mk_ongoing_code.x),
           mk_ethnicflag = ifelse(year == 2022, mk_ethnicflag.y, mk_ethnicflag.x),
           mk_ongoing = ifelse(year == 2022, mk_ongoing.y, mk_ongoing.x),
           mk_onset = ifelse(year == 2022, mk_onset.y, mk_onset.x),
           mk_offset = ifelse(year == 2022, mk_offset.y, mk_offset.x),
           widetargeting = ifelse(year == 2022, widetargeting.y, widetargeting.x),
           narrowtargeting = ifelse(year == 2022, narrowtargeting.y, narrowtargeting.x)) %>% 
    select(-datastartyear.y,-datastartyear.x, -dataendyear.y,-dataendyear.x,-mk_onset_count.y, -mk_onset_count.x, -mk_offset_count.y,-mk_offset_count.x,-mk_ongoing_count.y,-mk_ongoing_count.x, -mk_location_ever.y,-mk_location_ever.x,-mk_responsible_ever.y, -mk_responsible_ever.x,-mk_ongoing_widescope.y, -mk_ongoing_widescope.x,-mk_ongoing_code.y, -mk_ongoing_code.x,-mk_ethnicflag.y, -mk_ethnicflag.x,-mk_ongoing.y, -mk_ongoing.x,-mk_onset.y, -mk_onset.x,-mk_offset.y, -mk_offset.x,-widetargeting.y, -widetargeting.x,-narrowtargeting.y, -narrowtargeting.x)
    
```


# Append new data

This appends dat_new to dat_old, creating the full data-set

```{r}
# check to make sure you can append the new data

  if(length(setdiff(colnames(dat_old), colnames(dat_new)))>0){
    stop("different colnames")}


# append new data
  
  setcolorder(dat_new, colnames(dat_old))
  dat <- rbind(dat_old, dat_new, fill = T)
  
```


```{r, eval = F}
# save mass killing variables

  mkl_dat <- subset(dat, select = c("country", "ccode", "year", 
                                    "mk_onset", "mk_onset_count","mk_offset", "mk_offset_count",
                                    "mk_ongoing_count", "mk_location_ever","mk_responsible_ever",
                                    "mk_ongoing_code","mk_ongoing","widetargeting","narrowtargeting"))

data.table(mkl_dat)

  setkey(mkl_dat, country, year)
  
  fwrite(mkl_dat, file = "output/mkl_data.csv", row.names = F)
  
```


# Leads and Lags

In this section, I update the variables that involve leads and lags

## MK lead variables

```{r}

# order by year, country name, and ccode
    
  setkey(dat, year, country, ccode)

# create anymk lead variable, shifting by ccode
    
dat <- dat %>%
  group_by(ccode) %>%
  arrange(country, year) %>%
  mutate(
    mk.onset.1 = lead(mk_onset, default = NA),
    mk.onset.2 = lead(mk.onset.1, default = NA),
    mk.onset.3 = lead(mk.onset.1, n = 2, default = NA),
    mk.onset.2window = as.double((mk.onset.1 + mk.onset.2) > 0),
    mk.onset.3window = as.double((mk.onset.2window + mk.onset.3) > 0)
  ) %>%
  ungroup()
    
```

## Coup in the last 5 years


For countries that are created as a result of a coup, we want to ensure that our "coup.try.5yr" indicator shows that they had a coup in the first year of existence, though these are typically not tracked in the Powell & Tyne data. Therefore I turn coup.try.5yr on manually for these countries. This applies to Gabon (coup in 1964) and Djibouti (coup in 2000).

```{r}

# create a variable for the last year when an ccode had a coup

setDT(dat)
  
  dat[, last_coup_yr := ifelse(cou.any == 1, year, NA)]

# update last_coup_yr for coup-created countries
dat <- dat %>% 
  mutate(last_coup_yr = ifelse(country == "Gabon" & year %in% c(1964:1969),1964,last_coup_yr),
         last_coup_yr = ifelse(country == "Djibouti" & year %in% c(2000:2005),2000,last_coup_yr))

  dat[, last_coup_yr := na.locf(last_coup_yr, fromLast = F, na.rm = F), 
      by = "ccode"]
  dat[, last_coup_yr := ifelse(is.na(last_coup_yr), -Inf, last_coup_yr)]
  
# create variable for whether there was a coup in the last 5 years
  
  dat[, coup.try.5yr := as.double((year - last_coup_yr) <= 4)]

```

# Data Preparation


## Remake some variables

This section fills in some V-DEM variables by creating an adjusted version of the relevant WDI variable. This significantly cuts down on missingness.

```{r}
make.combined <- function(vdem.var, wdi.proxy){
  # Make an adjusted version of the wdi one to fit the tradeshare (VDEM) one:  
    # drop update year as we've carried forward some values to this year
  
    lm.adjust <- lm(as.formula(paste0(vdem.var, "~", wdi.proxy)), 
                    data = dat[year < update_years])
    
  # fill in update years with fitted values from regression on all years  
    
    dat[year %in% update_years, lm.adjust := coef(lm.adjust)[1] + 
          coef(lm.adjust)[2]*get(wdi.proxy)]
  
  # Where vdem var is missing, replace with the adjusted wdi proxy
    combined.var <- paste0(vdem.var, ".combined")
    dat[year %in% update_years, 
        (combined.var) := 
          ifelse(!is.na(get(vdem.var)), get(vdem.var), lm.adjust)]
    
  # summarize difference in missingness
    na.old <- sum(is.na(subset(dat[year %in% update_years], select = vdem.var)))
    na.new <- sum(is.na(subset(dat[year %in% update_years], select = combined.var))) 
  
    cat(paste0("In the update years we were missing ", na.old, 
           " observations for ", vdem.var))
    cat(paste0("\nUsing the WD indicator, ",  wdi.proxy, 
               ", as a proxy we are missing ",  na.new, " observations"))

    dat[, lm.adjust := NULL]
}    
```



```{r}
# create combined variable for tradeshare

  make.combined(vdem.var = "tradeshare.ln", wdi.proxy = "wdi.trade.ln.new")
  cat("\n")

# create combined variable for popsize.ln 

  make.combined(vdem.var = "popsize.ln", wdi.proxy = "wdi.popsize.ln")
  cat("\n")
  
    make.combined(vdem.var = "popsize", wdi.proxy = "wdi.popsize.new")
  cat("\n")

# create combined variable for GDP per capita growth

  make.combined(vdem.var = "gdppcgrowth", wdi.proxy = "wdi.gdppcgrow.new")
  
# create combined variable for infant mortality - hold on this vdem var is  not right
  # make.combined(vdem.var = "e_peinfmor", wdi.proxy = "wdi.imrate.new")
  #  cat("\n")

```


#Update Discrimpop, Ldiscrimpop, efindex and womeninparliament
EPR dataset - goes through 2020, carry forward (there is additional data avail online, check if it can be downloaded for future updates)
HIEF dataset - goes through 2013, carry forward
womeninparliament - goes through 2020, carry forward (there is additional data avail online, check if it can be downloaded for future updates)

#Fix old MK data to use in old model - Have to fix this data for Iraq in the old MK vars so that we can compare old to new model
```{r}
setDT(dat)
dat[country == "Iraq" & year == 2018, mkl.end := 1]
dat[country == "Iraq" & year == 2019, mkl.ongoing := 0]
dat[country == "Iraq" & year == 2020, mkl.ongoing := 0]
dat[country == "Iraq" & year == 2021, mkl.ongoing := 0]
dat[country == "Iraq" & year == 2022, mkl.ongoing := 0]
```

# Missingness



This section looks at remaining missingness and attempts to fill in values. 

```{r}
# select variables we would like to check missingness for (2023 update added "efindex", "discrimpop", "v2mecenefm_binary","v2csgender_binary", "mk_ongoing", "partyban.new.0", "partyban.new.1", "partyban.new.2", mk_location_ever","mk_responsible_ever", "widetargeting","narrowtargeting", "imr.sqrt.new" and removed polity2.fl.2","polity2.fl.3" and "durable.ln" as those are no longer used )


  predictornames <- c("anymk.ongoing","anymk.ever",
                      "reg.afr", "reg.eap", "reg.eur", "reg.mna", "reg.sca",
                      "countryage.ln", "popsize.ln.combined", "imr.sqrt.new",
                      "gdppcgrowth.combined", "ios.iccpr1","includesnonstate",
                      "minorityrule", "elf.ethnic", 
                      "battledeaths.ln", "candidaterestriction", "partyban", 
                      "judicialreform", "religiousfreedom", 
                      "pol_killing_approved", "freemove_men4", 
                      "freemove_women4", "freediscussion",
                      "social_inequality","even_civilrights","repress_civilsoc", 
                      "social_power_dist", "ses_power_dist",
                      "tradeshare.ln.combined", 
                      "coup.try.5yr",
                      "efindex", "discrimpop", "v2mecenefm_binary",
                      "mk_ongoing", "partyban.new.0", 
                      "partyban.new.1", "partyban.new.2",
                      "mk_location_ever","mk_responsible_ever",
                      "widetargeting","narrowtargeting", "wdi.gdppc.new")


# look for missingness in the update years for select variables

  dat.check <- subset(dat, year %in% update_years, 
                      select = c("country","year", predictornames))

# for each of the update years, count the number of NAs for each variable
# select variables with positive NA counts

  comp <- lapply(update_years, 
                 function(y) apply(dat.check[year == y], 
                                   2, function(x) sum(is.na(x))))
  na.count <- unlist(comp)
  years <- rep(update_years, each = ncol(dat.check))
  check <- data.table(na.count, years, "var"=names(na.count))
  check <- check[na.count > 0]

# visualize missingness
  
   (missing1 <- ggplot(check) + 
      geom_bar(aes(y = na.count, x = var, fill = factor(years)), 
               stat = "identity")+ 
      theme(axis.text.x = element_text(angle = 45, hjust = 1)) + 
      labs(x = "Missing Variables", y = "Number of Observations Missing", 
           fill = "Year", 
           title = "Number of Countries with Missing Values Before Fixing"))

```

## Fix missingness

This chunk defines functions to look at patterns in NAs and carry forward values.

```{r}

# function to look at patterns in NAs over time for each country with missing values

  look_na <- function(variable, data = dat){
    missing <- data[is.na(get(variable)) & 
                     year %in% update_years, c("year", "country")]
    countries <- unique(missing$country)
    if(length(countries) == 0){
      look <- "No missing values"
    }else{
      look <-lapply(1:length(countries), function(x){
        out <- subset(data, country == countries[x],
                      select = c("year", variable)) 
        colnames(out) <- c("year", countries[x])
        out})
      
      # merge all data.tables in the list by year
        look <- Reduce(function(...) merge(..., all = T, by = "year"), look)
        colnames(look) <- c("year", as.character(countries))
        look <- look[order(look$year, decreasing = TRUE), ]
      }
    look
  }


# function to create blank spreadsheets to fill out 
  
  fill.tab <- function(var){
    out <- subset(dat[is.na(get(var)) & year %in% update_years], 
                  select = c("country", var))
    out[, cia.factbook.est := NA]
    out[, est.year := NA]
    fwrite(out, file = paste0("Make-new-base-data/input/temp/missing/", var, ".csv"))
  }
  

# function to look at last present value for countries missing data in update years
  
  last.present <- function(var){
    look.na.long <- melt(look.na.vars[[var]], id.vars = "year", 
                         variable.name = "country")
    setorder(look.na.long, country, year)
    look.na.long[, value_last := shift(value), by = country]
    look.na.long[, same_as_last := value == value_last]
    look.na.long[!is.na(value) & is.na(value_last), same_as_last := F]
    look.na.long[!is.na(same_as_last) & same_as_last == F, 
                 .("carried_value" = value[which.max(year)], 
                   "carry_from" = max(year)), by = .(country)]

  }

```


```{r}
# for each of the variables that we have countries with missing values
# look at the values in the last few years for that country
#creates blank spreadsheets to fill out

  look.na.vars <- lapply(c(check$var, "imr.sqrt.new"), look_na)
  names(look.na.vars) <- c(check$var, "imr.sqrt.new")

  if(first.pass == T){
    dir.create("Make-new-base-data/input/temp/missing")  
    lapply(check$var, fill.tab)
    fill.tab("imr.sqrt.new")
    }  
```


### Candidate Restriction

```{r}      
# candidate restriction

  head(look.na.vars$candidaterestriction) 

#fill in prior years that were carried forward (they were removed as a result of new vdem merge for 2022 update)
dat[country == "Jamaica" & year == 2019, candidaterestriction := 1]
dat[country == "Jamaica" & year == 2020, candidaterestriction := 1]
dat[country == "Jamaica" & year == 2021, candidaterestriction := 1]
dat[country == "Jamaica" & year == 2022, candidaterestriction := 1]

  
# just carry forward this value 
  
  carry_forward("candidaterestriction")

```

For the variables addressed below, I read in values from the CIA Factbook (which were entered manually into the spreadsheet created above), and carry those values forward if they're more recent than the last updated version of the variable.

In the tables presented for each variable, I show the value that is filled in for that country in column 2. This is either the "carried_value", carried from the year listed under "carry_from", or it is the "cia.factbook.est".

### Infant Mortality

#10/20 - holding on  this because we are interpolating

Countries filled in for 2022:
- Kosovo

```{r, eval = FALSE}

 # head(look.na.vars$imr.sqrt)

# read in CIA Factbook estimates

#  cia.factbook <- fread("input/temp/missing/imr.sqrt.csv")
 # cia.factbook[, cia.factbook.est := sqrt(cia.factbook.est)]
  
  
  # subtract one from the carry_from value because there is a lag
  # in updating the infant mortality rate
  
#  comp <- merge(cia.factbook, last.present("imr.sqrt"), 
                by = "country")
 # comp[, imr.sqrt := ifelse(carry_from >= est.year, 
                                        carried_value, cia.factbook.est)]
#  kable(comp)
  
# replace missing

#  for(i in 1:nrow(comp)){
 #   country.fill <- comp[i, ]
  #  dat[country ==  country.fill$country & year == update_years, 
   #     imr.sqrt := country.fill$imr.sqrt]
  }
  


```

### Population size 

Countries filled in for 2022:
- Taiwan


```{r} 

  head(look.na.vars$popsize.ln.combined)


# read in CIA Factbook estimates

  cia.factbook <- fread("input/temp/missing/popsize.ln.combined.csv")
  
  comp <- merge(cia.factbook, last.present("popsize.ln.combined"), 
                by = "country")
    
  comp[, popsize.ln.combined := ifelse(carry_from >= est.year, 
                                       carried_value, log(cia.factbook.est))]
  comp[, popsize.combined := ifelse(carry_from >= est.year, 
                                        carried_value, cia.factbook.est)]
  kable(comp)
  
# replace missing
  
  setDT(dat)

  for(i in 1:nrow(comp)){
    country.fill <- comp[i, ]
    dat[country ==  country.fill$country & year == update_years, 
        popsize.ln.combined := country.fill$popsize.ln.combined]
  }

  for(i in 1:nrow(comp)){
    country.fill <- comp[i, ]
    dat[country ==  country.fill$country & year == update_years, 
        popsize.combined := country.fill$popsize.combined]
  }
```


### GDP Per capita growth

In 2022, this variable is missing for the following countries: 

```{r, echo = F}

  head(look.na.vars$gdppcgrowth.combined)

missing_22 <- c("Afghanistan", "Bhutan","Cuba", "Eritrea","Korea, People's Republic of", "Kosovo", "Lebanon", "South Sudan", "Syria", "Taiwan", "Turkmenistan", "Venezuela", "Yemen (unified)", "Yugoslavia")
cat(paste(paste(missing_22, collapse = ", "), "\n"))
```

The following countries are newly missing, and have more recent estimates from the WDI than the CIA factbook, so we carry old WDI values forward (none applicable)

```{r}
missing_23 = colnames(look.na.vars$gdppcgrowth.combined)[-1]
cat(paste(paste(setdiff(missing_23, missing_22), collapse = ", "), "\n"))
```


```{r}

# read in CIA Factbook estimates
# Called "Real GDP growth rate" IN FACTBOOK

  cia.factbook <- fread("input/temp/missing/gdppcgrowth.combined.csv")
  comp <- merge(cia.factbook, last.present("gdppcgrowth.combined"), 
                by = "country")
  comp[, cia.factbook.est := cia.factbook.est/100]

  
  comp[, gdppcgrowth.combined := ifelse(carry_from >= est.year, 
                                        carried_value, cia.factbook.est)]
  comp[, source := factor(carry_from >= est.year & as.character(round(carried_value, 4)) != as.character(cia.factbook.est), levels = c("TRUE", "FALSE"), labels = c("carry_WDI", "CIA_Factbook"))]

  kable(comp[order(source)], digits = 3)
  
# replace missing
  
  setDT(dat)

  for(i in 1:nrow(comp)){
    country.fill <- comp[i, ]
    dat[country ==  country.fill$country & year == update_years, 
        gdppcgrowth.combined := country.fill$gdppcgrowth.combined]
  }

```



### tradeshare.ln.combined

#### First fill in GDP per capita
Filled in :
- Eritrea, North Korea, South Sudan, Taiwan, Venezuela, Yemen

```{r}

  head(look.na.vars$wdi.gdppc.new)

  cia.factbook <- fread("input/temp/missing/wdi.gdppc.new.csv")
  comp <- merge(cia.factbook, last.present("wdi.gdppc.new"), 
                by = "country", all.x = T)

  comp[, wdi.gdppc.new := ifelse((est.year < (carry_from-1) & !is.na(carry_from))|
                                   is.na(cia.factbook.est), carried_value, 
                                 cia.factbook.est)]
  comp[, source := factor(carry_from-1 > est.year & as.character(round(carried_value, 4)) != as.character(cia.factbook.est), levels = c("TRUE", "FALSE"), labels = c("carry_WDI", "CIA_Factbook"))]
  kable(comp[order(source)], digits = 3)
  
  # replace missing
  
  setDT(dat)

  for(i in 1:nrow(comp)){
    country.fill <- comp[i, ]
    dat[country ==  country.fill$country & year == update_years, 
        wdi.gdppc.new := country.fill$wdi.gdppc.new]
  }
```


#### Calculate Tradeshare

In the 2021 update, we used CIA Factbook estimates for the following countries: 

```{r, echo = F}
missing_21 <- c("Afghanistan", "Guyana", "Laos", "Malawi", "Taiwan")
missing_22 <- colnames(look.na.vars$tradeshare.ln.combined)[-1]
cat(paste(paste(missing_21, collapse = ", "), "\n"))
```

Many of these countries still have missing values during the 2022 update, so we continue to use CIA Factbook values for them. The following countries are newly missing tradeshare data in 2022:

```{r, echo = F}
 cat(paste(paste(intersect(missing_22, missing_21), collapse = ", "), "\n"))
cat(paste(paste(setdiff(missing_22, missing_21), collapse = ", "), "\n"))
```



```{r}
# read in CIA Factbook estimates - be sure to enter in billions 

  head(look.na.vars$tradeshare.ln.combined)

  cia.factbook <- fread("input/temp/missing/tradeshare.ln.combined.csv")
  comp <- merge(cia.factbook, last.present("tradeshare.ln.combined"), 
                by = "country")
  other.vars <- dat[year == update_years, 
                    c("wdi.gdppc.new", "popsize.combined", "country")]
  comp <- merge(comp, other.vars, by= "country", all.x = T)
  
    
  # The VDEM imports and exports are in millions of USDs, then x10^6. 
  # tradeshare = (imports + exports (in trillions))/(gdppc*popsize)
  # The CIA factbook data is in billions, so needs to be multiplied by 1e+9
  comp[, cia.factbook.est := log(1e+9*cia.factbook.est/(wdi.gdppc.new*popsize.combined))]


  
    comp[, tradeshare.ln.combined := 
           ifelse(est.year <= carry_from, carried_value, cia.factbook.est)]
    comp[, source := factor(carry_from >= est.year & as.character(round(carried_value, 4)) != as.character(cia.factbook.est), levels = c("TRUE", "FALSE"), labels = c("carry_WDI", "CIA_Factbook"))]
    comp[country %in% intersect(missing_22, missing_21), source := "CIA_Factbook"]
    kable(subset(comp[order(source)], select = c("country", "tradeshare.ln.combined",
                                  "cia.factbook.est", "est.year", 
                                  "carried_value", "carry_from", "source")))

# replace missing

  for(i in 1:nrow(comp)){
    country.fill <- comp[i, ]
    dat[country ==  country.fill$country & year == update_years, 
        tradeshare.ln.combined := country.fill$tradeshare.ln.combined]
  }
    

  # carry_forward("tradeshare.ln.combined")
  
```



## Check that all is filled in

```{r}
# repeat to make sure there are no more missing
    
  look.na.vars <- lapply(check$var, function(x) 
    look_na(x, dat = dat[year>=1945 & country!="Taiwan"]))
  names(look.na.vars) <- check$var
  look.na.vars
  

```



#Export the new prepared predictors file:
```{r}

fwrite(dat, file = paste0("output/prepared", update_years, "predictors-", Sys.Date(), "-2", ".csv"))

```

